---
title: "Analyses for honest signals MS"
output: github_document
---

## Analyses

Question 1: Is flower size variation consistent between the field and the greenhouse?

*Approach*: Comparing population means between the field and the greenhouse. Are we trying to compare each individual population field to greenhouse (e.g. two sample t-tests), or a trend across the populations field vs. greenhouse (e.g. regression of population means between field and greenhouse)?

Question 2: Are size and scent honest signals for nectar rewards?

*Approach*: Correlation between size and rewards, and between scent and rewards. First do a regression of all plants, with plant nested within population as a random effect. Then do regressions for each population.

  
### Packages and start up code:

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
library(knitr)
library(lmerTest)
library(emmeans)
library(ggpubr)
library(plotrix)
```

### Q1: Is there population differentiation in size in the field and the greenhouse?

*Greenhouse*
Loading in the data, filtering out the NAs, and selecting the populations with 8 or more plants
```{r, warning=FALSE, message=FALSE}
nectar <- read_excel("nectar protocol.xlsx")

nectar_work <- nectar %>% filter(!is.na(Population)) %>%  filter(!is.na(nectar_μl)) %>% filter(!is.na(Corola_area_mm2))

nectar_work8 <- nectar_work %>% group_by(Population) %>% filter (length(unique(Plant_id))>=8)

```
Analyzing population differences

```{r}
size_mod<- lmer(Corola_area_mm2~Population+(1|Plant_id), data = nectar_work8)

anova(size_mod)

#emmeans(size_mod, pairwise~Population)
emms<-emmeans(size_mod, "Population")
#pwpm(emms)
plot(emms, comparisons = TRUE)

```

*Field*

```{r, message=FALSE,warning=FALSE}
field <- read_csv("field_data.csv")
field_work <- field %>% filter(!is.na(FlowerSize))

field_work %>% group_by(Population) %>% summarize(N=length(Id))

size_mod<- lm(FlowerSize~Population, data = field_work)

anova(size_mod)

emms<-emmeans(size_mod, pairwise~Population)
plot(emms, comparisons = TRUE)

```

*comparing the field and the greenhouse*

```{r}
Pops <- c("G3", "I10", "I15", "I16", "I17", "I4","It10", "It15", "It16", "It17", "It4")

field_combine <- field_work %>% filter(Population %in% Pops) %>%  group_by(Population) %>% summarize(FlSizeMean = mean(FlowerSize), FlSizeSE=std.error(FlowerSize)) %>% mutate(Source="Field")

gh_combine <- nectar_work8 %>% filter(Population %in% Pops) %>%  group_by(Population) %>% summarize(FlSizeMean = mean(Corola_area_mm2), FlSizeSE=std.error(Corola_area_mm2)) %>% mutate(Source="Greenhouse")

combine <- rbind(field_combine, gh_combine)
combine[8:12,1]<-c("I10", "I15", "I16", "I17", "I4")

combine_wide <- combine %>% pivot_wider(names_from=Source, values_from=c(FlSizeMean,FlSizeSE))

combine_mod <- lm(FlSizeMean_Greenhouse~FlSizeMean_Field, data=combine_wide)

ggplot(aes(x=FlSizeMean_Field, y=FlSizeMean_Greenhouse), data=combine_wide)+geom_point()+geom_errorbar(aes(ymin=FlSizeMean_Greenhouse-FlSizeSE_Greenhouse, ymax=FlSizeMean_Greenhouse+FlSizeSE_Greenhouse))+geom_errorbarh(aes(xmin=FlSizeMean_Field-FlSizeSE_Field, xmax=FlSizeMean_Field+FlSizeSE_Field))+xlab("Flower size in natural populations")+ylab("Flower size in greenhouse")+theme_classic()


```



### Q2: Are size and scent honest signals for nectar rewards?

#### scent work

This is an object with average nectar values per plant for those populations with sufficient plants. We can merge these data with the scent data.

```{r, message=FALSE,warning=FALSE}
nectar <- read_excel("nectar protocol.xlsx")

nectar_work <- nectar %>% filter(!is.na(Population)) %>%  filter(!is.na(nectar_μl)) %>% filter(!is.na(Corola_area_mm2))

nectar_work8 <- nectar_work %>% group_by(Population) %>% filter (length(unique(Plant_id))>=8)

nectar_8_avg <- nectar_work8 %>% group_by(Population, Plant_id) %>% summarize(Mean_nectar=mean(nectar_μl))

#add scent into this df

#nectar_model<-lmer(nectar_μl~scent+(1|Population), data=nectar_8_avg)

#par(mfrow=c(1,2))
#hist(resid(nectar_model))
#plot(predict(nectar_model),resid(nectar_model)) ; abline(h=0)

```

```{r, warning=FALSE, message=FALSE}
#ggplot(aes(x=scent, y=nectar_μl), #data=nectar_8_avg)+geom_point(aes(color=Population))+geom_smooth(method="lm",color="black")+theme_classic()+xlab(expression(paste("Total scent emission (ng", h^-1, flower^-1, ")"))) + ylab(expression(paste("Nectar volume (", mu,"l)")))
```

#### nectar work
```{r, warning=FALSE, message=FALSE}
nectar <- read_excel("nectar protocol.xlsx")
```

#### Summarizing and subsetting the dataset 

*summarizing the NAs*

```{r}
#kable(nectar %>% filter(is.na(Population)) %>% group_by(Plant_id) %>% summarise(N=length(unique(flower_id))))
```

*removing flowers where we don't have the population and we don't have nectar and/or flower size*

```{r}
nectar_work <- nectar %>% filter(!is.na(Population)) %>%  filter(!is.na(nectar_μl)) %>% filter(!is.na(Corola_area_mm2))
```


*summarizing how many flowers and plants we have data for*

```{r, warning=FALSE, message=FALSE}

nectar_work_sum <- nectar_work %>% group_by(Population) %>%  summarise(N_plants=length(unique(Plant_id)), N_flowers=length(flower_id))
kable(nectar_work_sum)
```


*making a list of populations with 8 or more plants sampled*

```{r}
nectar_list <- nectar_work_sum %>% filter(N_plants >= 8)
```


*filtering the dataset to only contain those populations*

```{r}
nectar_work8 <- nectar_work %>% group_by(Population) %>% filter (length(unique(Plant_id))>=8)
```

*adding in classification of mating systems*

```{r}
nectar_work8$Mating_system <- if_else(str_detect(nectar_work8$Population,"^G|^I"), "SI", "SC")
```


### Analysis: nectar vs. size, all populations

```{r}
nectar_model<-lmer(nectar_μl~Corola_area_mm2+(1|Population/Plant_id), data=nectar_work8)

par(mfrow=c(1,2))
hist(resid(nectar_model))
plot(predict(nectar_model),resid(nectar_model)) ; abline(h=0)
```

The residuals get a little better with a square-root transformation, but it's not a huge difference.

```{r, warning=FALSE, message=FALSE}


ggplot(aes(x=Corola_area_mm2, y=nectar_μl), data=nectar_work8)+geom_point(aes(color=Population))+geom_smooth( method="lm",color="black")+theme_classic()+xlab(expression(paste("Corolla area (", mm^2 ,")"))) + ylab(expression(paste("Nectar volume (", mu,"l)")))
```

I think this works for now, but if we use all of this data, it's probably not effective to show this many populations in colors. Maybe we don't need to color the points by population here if we do a population by population analysis. Another option could be coloring the points by mating system, or region.

## Analysis, nectar vs. size with mating system added

```{r}

nectar_work8$Mating_system<-factor(nectar_work8$Mating_system, levels=c("SI", "SC"))
nectar_model<-lmer(nectar_μl~Corola_area_mm2*Mating_system+(1|Population/Plant_id), data=nectar_work8)

par(mfrow=c(1,2))
hist(resid(nectar_model))
plot(predict(nectar_model),resid(nectar_model)) ; abline(h=0)

anova(nectar_model)
emmeans(nectar_model, "Mating_system")

ggplot(aes(x=Corola_area_mm2, y=nectar_μl), data=nectar_work8)+geom_point(aes(color=Mating_system, shape=Mating_system), stroke=1.2)+geom_smooth(aes(color=Mating_system), method="lm")+geom_smooth(aes(color="All", shape="All"), method="lm", linetype=2, show.legend=TRUE)+theme_classic()+xlab(expression(paste("Corolla area (", mm^2 ,")"))) + ylab(expression(paste("Nectar volume (", mu,"l)")))+scale_color_discrete(name="Mating system", breaks=c("All", "SC", "SI"), labels=c("All", "Self-compatible", "Self-incompatible"))+scale_shape_manual(name="Mating system",values=c(NA, 0,1),breaks=c("All", "SC", "SI"), labels=c("All", "Self-compatible", "Self-incompatible"))

#size vs. mating system

size_model<-lmer(Corola_area_mm2~Mating_system+(1|Population/Plant_id), data=nectar_work8)

par(mfrow=c(1,2))
hist(resid(size_model))
plot(predict(size_model),resid(size_model)) ; abline(h=0)

anova(size_model)


ggplot(aes(x=Mating_system, y=Corola_area_mm2), data=nectar_work8) + geom_boxplot()+theme_classic()+ylab(expression(paste("Corolla area (", mm^2 ,")"))) + xlab("Mating system")
```


## Analysis: nectar vs. size, population by population

```{r, warning=FALSE, message=FALSE}

#writing a function to generate a model summary

f <- function(df) summary(lmer(nectar_μl ~ Corola_area_mm2 +(1|Plant_id), data = df))

#running this model on the data subset by population

v <- lapply(split(nectar_work8, nectar_work8$Population), f)

# making a list of the populations and a blank dataframe that will store the results 

Pop=sort(as.factor(c(unique(nectar_work8$Population),unique(nectar_work8$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

#extracting the coefficients from each model

for (i in 1:29){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

# adding the population labels to these results

results<-cbind(results, Pop)

#cleaning up the results table to only contain the slope estimates; adding a significance column
results<-results %>% tibble::rownames_to_column("X") 
results <- results %>% filter(str_detect(X, "Corola"))  %>% mutate(Sig=if_else(`Pr(>|t|)`< 0.05, "Yes", "No")) %>% select(-X) %>% relocate(Pop) 

kable(results)

#creating an object of just the significant ones, to then create an object with only the data from populations where there is a significant correlation, so that the plot only has lines drawn for the significant relationships
results.sig<-results %>% filter(Sig=="Yes")
lines<-nectar_work8 %>% filter(Population %in% results.sig$Pop)

##plotting the population level results
ggplot(aes(x=Corola_area_mm2, y=nectar_μl, color=Population), data=nectar_work8)+geom_point()+
  facet_wrap(~Population, scales="free")+theme_classic()+xlab(expression(paste("Corolla area (", mm^2 ,")"))) + ylab(expression(paste("Nectar volume (", mu,"l)")))+geom_smooth(data=lines, method="lm", color="black")


```




