---
title: "Seed estimates and pollen limitation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

In this document, I create the regressions to estimate seed set in the fruits that were open in Italy. I then use these seed set estimates, along with the seed counts, to calculate the extent of pollen limitation in all populations.

```{r, warning=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(lme4)
library(lmerTest)
library(knitr)
library(emmeans)
library(ggpubr)
library(plotrix)
library(magicfor)
```

Loading in the data and filtering out rows with missing values. Created two objects, one with the Italian populations and one with the Greek populations.
```{r, warning=FALSE,message=FALSE}
seeds_E <- read_excel("Seed_estimate.xlsx")


seeds_E <- seeds_E %>% filter(!is.na(Missing_seeds)) %>% filter(!is.na(Treatment))

seeds_I <- seeds_E %>% filter(!str_detect(Population,"^G")) 
seeds_G <- seeds_E %>% filter(str_detect(Population,"^G")) 
```

The number of non HP and C fruits used in each population for the regressions

```{r, warning=FALSE,message=FALSE}
kable(seeds_I %>% filter(Treatment != "c" & Treatment != "hp" & fruit_length >= 5) %>% group_by(Population, Treatment) %>% summarize(N=length(Missing_seeds)))


```


The percentages of fruits (from the control or hp treatments) that were missing seeds

```{r, warning=FALSE,message=FALSE}
seeds_I %>% filter(Treatment == "c" | Treatment == "hp")  %>%  group_by(Population) %>% summarize(N_missing=sum(Missing_seeds), N_total=length(Missing_seeds)) %>% mutate(Per_missing=100*(N_missing/N_total))
```

The number of fruits (broken out by treatments) that were intact

```{r, warning=FALSE,message=FALSE}
kable(seeds_I %>% filter(Treatment == "c" | Treatment == "hp" & fruit_length >= 5)  %>% group_by(Population, Treatment) %>% summarize(N_intact=length(Missing_seeds)-sum(Missing_seeds)))
```



Okay so for a given population, we want to take the intact fruits where fruit length was measured and was greater than or equal to 5, and run a regression of fruit length vs. seed set with plant ID as a random effect

```{r, warning=FALSE,message=FALSE}
#creating a new object to contain only fruits that weren't missing seeds and that had a length greater than or equal to 5
seeds_est <- seeds_I %>% filter(Missing_seeds==0 & fruit_length >= 5 & Seeds_counted!="NA")

#creating a function to return the summary of the model we want to apply to all the population subsets
f <- function(df) summary(lmer(Seeds_counted ~ fruit_length +(1|Plant_ID), data = df))

#subsetting the data by population and applying the function
v <- lapply(split(seeds_est, seeds_est$Population), f)

#creating the list of populations and the empty data frame for the results
Pop=sort(as.factor(c(unique(seeds_est$Population),unique(seeds_est$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

##extracting the coefficients from each model
for (i in 1:6){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

#putting the results together with population labels
results<-cbind(results, Pop)
results<-results %>% tibble::rownames_to_column("Coefficient") 
kable(results)

#Using these regressions to calculate the estimated number of seeds for fruits that had missing seeds
#My approach was to loop over the populations, subset the data by that population, and then use the regression parameters for that population to create a column that contains either the counted seed number (if no seeds were missing), or the estimated seed number if seeds were missing
magic_for()
for (pop in unique(Pop)) {
  a <- seeds_I %>% filter(Population==pop) 
  aa <- results %>% filter(Pop==pop)
  b <- if_else(a$Missing_seeds==1, (a$fruit_length*aa[2,2])+aa[1,2], a$Seeds_counted)
  put(b)
}

#get the result of the loop and unlist it to make it a column of the seeds_I object
z<-magic_result()
magic_free()
seeds_I$estimates<-unlist(z)

```


Here we are extracting the predicted values from each model fit, to use in plotting

```{r, warning=FALSE,message=FALSE}



pop_list <- unique(seeds_est$Population)

magic_for()
for (pop in pop_list) {
  a <- seeds_est %>% filter(Population==pop) 
  b <- lmer(Seeds_counted ~ fruit_length +(1|Plant_ID), data = a)
  c <- predict(b, type= "response")
  put(c)
  
}
#extracting the results and putting them into a column
z<-magic_result()
magic_free()
seeds_est$pred<-unlist(z)

#plotting the regressions for each population
ggplot(aes(x=fruit_length),data=seeds_est)+geom_point(aes(y=Seeds_counted)) + geom_smooth(aes(y=pred), method="lm")+facet_wrap(~Population)+theme_classic()+xlab("Fruit length (mm)")+ylab("Seed number")

```

Now we need to take the `seeds_I$estimates` and put them into `seeds_E`, to use for our calculations for pollen limitation for all populations

```{r, warning=FALSE,message=FALSE}
seeds_G$estimates <- seeds_G$Seeds_counted
seeds_work <- rbind(seeds_G, seeds_I)
```

Working with only the `hp` and `c` fruits, and calculating how many plants had zeros
```{r, warning=FALSE,message=FALSE}
seeds_work <- seeds_work %>% filter(Treatment == "hp" | Treatment == "c")

#looking at number of plants that had zeros in the hp fruits, to see if we really need to drop plants 
zees<-seeds_work %>% group_by(Population, Plant_ID) %>% filter(Treatment== "hp") %>% summarize(zeros=sum(estimates==0), nonzeroes=sum(estimates!=0))

zees2<-seeds_work %>% group_by(Population, Plant_ID, Treatment) %>%  summarize(zeros=sum(estimates==0), nonzeroes=sum(estimates!=0)) %>% pivot_wider(names_from=Treatment, values_from=c(zeros, nonzeroes))


#for now, we will only analyze plants where there were fewer than 2 zero-seed hand-pollinated fruits, following Sotiria's criteria
#this is a list of the plants had less than 2 zeros in the hp fruits
keep_list <- zees2 %>% filter(zeros_hp <2) %>% mutate(ID=paste(Population, Plant_ID))

#making an object with only those plants
seeds_work2<- seeds_work %>% mutate(ID=paste(Population, Plant_ID)) %>% filter(ID %in% keep_list$ID) %>% mutate(Treatment=factor(Treatment, levels=c("hp", "c")))

#this tells us sample sizes
seeds_work2 %>% group_by(Population) %>% summarize(N=length(unique(Plant_ID)))

```


Run an lmer for each population

```{r, warning=FALSE, message=FALSE}
#creating a function that extracts the emmeans for models
f <- function(df) as.data.frame(emmeans(lmer(estimates ~ Treatment +(1|Plant_ID), data = df), "Treatment"))
#applying this function to the dataset subset by population
v <- lapply(split(seeds_work2, seeds_work2$Population), f)

#creating a function that extracts the pairwise comparisons of emmeans for models
f2 <- function(df) as.data.frame(pairs(emmeans(lmer(estimates ~ Treatment +(1|Plant_ID), data = df), "Treatment")))
#applying this function to the dataset subset by population
v3 <- lapply(split(seeds_work2, seeds_work2$Population), f2)

#making the contrasts object more usable, adding significance column
v4 <-do.call(rbind.data.frame, v3)
v4$Pop<-rownames(v4)
rownames(v4)<-c()
v4$sig<-ifelse(v4$p.value<0.05, "*", "")

#making the emmeans object more usable
v2<-do.call(rbind.data.frame, v)
Pop=sort(as.factor(c(unique(seeds_work2$Population),unique(seeds_work2$Population))))
v2<-data.frame(v2, Pop, row.names = c())

#printing these tables
kable(v2)
kable(v4)

#plotting the pollen limitation results
ggplot(aes(x=Treatment, y=emmean), data=v2) + facet_wrap(~Pop) +geom_point()+geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=0.2) +ylab("Seed set") + theme_classic()+geom_text(aes(x=1.5, y=20, label=sig),data=v4, size=6 )

```