---
title: "Seed estimates and pollen limitation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

In this document, I create the regressions to estimate seed set in the fruits that were open in Italy. I then use these seed set estimates, along with the seed counts, to calculate the extent of pollen limitation in all populations.

```{r, warning=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(lme4)
library(lmerTest)
library(knitr)
library(emmeans)
library(ggpubr)
library(plotrix)
library(magicfor)
library(ggrepel)
```

Loading in the data and filtering out rows with missing values. Created two objects, one with the Italian populations and one with the Greek populations.
```{r, warning=FALSE,message=FALSE}
field <- read_csv("honest_signals/field_data.csv")

seeds_E <- read_excel("honest_signals/Seed_estimate.xlsx")


seeds_E <- seeds_E %>% filter(!is.na(Missing_seeds)) %>% filter(!is.na(Treatment))

seeds_I <- seeds_E %>% filter(!str_detect(Population,"^G")) 
seeds_G <- seeds_E %>% filter(str_detect(Population,"^G")) 
```

The number of non HP and C fruits used in each population for the regressions

```{r, warning=FALSE,message=FALSE}
kable(seeds_I %>% filter(Treatment != "c" & Treatment != "hp" & fruit_length >= 5) %>% group_by(Population, Treatment) %>% summarize(N=length(Missing_seeds)))


```


The percentages of fruits (from the control or hp treatments) that were missing seeds

```{r, warning=FALSE,message=FALSE}
seeds_I %>% filter(Treatment == "c" | Treatment == "hp")  %>%  group_by(Population) %>% summarize(N_missing=sum(Missing_seeds), N_total=length(Missing_seeds)) %>% mutate(Per_missing=100*(N_missing/N_total))
```

The number of fruits (broken out by treatments) that were intact

```{r, warning=FALSE,message=FALSE}
kable(seeds_I %>% filter(Treatment == "c" | Treatment == "hp" & fruit_length >= 5)  %>% group_by(Population, Treatment) %>% summarize(N_intact=length(Missing_seeds)-sum(Missing_seeds)))
```



Okay so for a given population, we want to take the intact fruits where fruit length was measured and was greater than or equal to 5, and run a regression of fruit length vs. seed set with plant ID as a random effect

```{r, warning=FALSE,message=FALSE}
#creating a new object to contain only fruits that weren't missing seeds and that had a length greater than or equal to 5
seeds_est <- seeds_I %>% filter(Missing_seeds==0 & fruit_length >= 5 & Seeds_counted!="NA") %>% 
  mutate(Plant_ID_new = if_else(Treatment=="n", Plant_ID+400,Plant_ID)) %>% 
  mutate(Plant_ID_final = if_else(Treatment=="e", Plant_ID+500, Plant_ID_new))

#creating a function to return the summary of the model we want to apply to all the population subsets
f <- function(df) summary(lmer(Seeds_counted ~ fruit_length +(1|Plant_ID_final), data = df))

#subsetting the data by population and applying the function
v <- lapply(split(seeds_est, seeds_est$Population), f)

#creating the list of populations and the empty data frame for the results
Pop=sort(as.factor(c(unique(seeds_est$Population),unique(seeds_est$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

##extracting the coefficients from each model
for (i in 1:6){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

#putting the results together with population labels
results<-cbind(results, Pop)
results<-results %>% tibble::rownames_to_column("Coefficient") 
kable(results)

#Using these regressions to calculate the estimated number of seeds for fruits that had missing seeds
#My approach was to loop over the populations, subset the data by that population, and then use the regression parameters for that population to create a column that contains either the counted seed number (if no seeds were missing), or the estimated seed number if seeds were missing
magic_for()
for (pop in unique(Pop)) {
  a <- seeds_I %>% filter(Population==pop) 
  aa <- results %>% filter(Pop==pop)
  b <- if_else(a$Missing_seeds==1, (a$fruit_length*aa[2,2])+aa[1,2], a$Seeds_counted)
  put(b)
}

#get the result of the loop and unlist it to make it a column of the seeds_I object
z<-magic_result()
magic_free()
seeds_I$estimates<-unlist(z)

```


Here we are extracting the predicted values from each model fit, to use in plotting

```{r, warning=FALSE,message=FALSE}



pop_list <- unique(seeds_est$Population)

magic_for()
for (pop in pop_list) {
  a <- seeds_est %>% filter(Population==pop) 
  b <- lmer(Seeds_counted ~ fruit_length +(1|Plant_ID), data = a)
  c <- predict(b, type= "response")
  put(c)
  
}
#extracting the results and putting them into a column
z<-magic_result()
magic_free()
seeds_est$pred<-unlist(z)

#plotting the regressions for each population
ggplot(aes(x=fruit_length),data=seeds_est)+geom_point(aes(y=Seeds_counted)) + geom_smooth(aes(y=pred), method="lm")+facet_wrap(~Population)+theme_classic()+xlab("Fruit length (mm)")+ylab("Seed number")

```

Now we need to take the `seeds_I$estimates` and put them into `seeds_E`, to use for our calculations for pollen limitation for all populations

```{r, warning=FALSE,message=FALSE}
seeds_G$estimates <- seeds_G$Seeds_counted
seeds_work <- rbind(seeds_G, seeds_I)
```

Working with only the `hp` and `c` fruits, and calculating how many plants had zeros
```{r, warning=FALSE,message=FALSE}
seeds_work <- seeds_work %>% filter(Treatment == "hp" | Treatment == "c")

#looking at number of plants that had zeros in the hp fruits, to see if we really need to drop plants 
zees<-seeds_work %>% group_by(Population, Plant_ID) %>% filter(Treatment== "hp") %>% summarize(zeros=sum(estimates==0), nonzeroes=sum(estimates!=0))

zees2<-seeds_work %>% group_by(Population, Plant_ID, Treatment) %>%  summarize(zeros=sum(estimates==0), nonzeroes=sum(estimates!=0)) %>% pivot_wider(names_from=Treatment, values_from=c(zeros, nonzeroes))


#for now, we will only analyze plants where there were fewer than 2 zero-seed hand-pollinated fruits, following Sotiria's criteria
#this is a list of the plants had less than 2 zeros in the hp fruits
keep_list <- zees2 %>% filter(zeros_hp <2) %>% mutate(ID=paste(Population, Plant_ID))

#making an object with only those plants
seeds_work2<- seeds_work %>% mutate(ID=paste(Population, Plant_ID)) %>% filter(ID %in% keep_list$ID) %>% mutate(Treatment=factor(Treatment, levels=c("hp", "c")))

#this tells us sample sizes
seeds_work2 %>% group_by(Population) %>% summarize(N=length(unique(Plant_ID)))

seeds_work4<- seeds_work2 %>% mutate(estimates=if_else(estimates<0, 0, estimates)) %>% mutate(estimates=if_else(estimates==333, 33, estimates)) %>% mutate(estimates=if_else(is.na(estimates), Seeds_counted, estimates)) %>% drop_na(estimates)

oppsel <- seeds_work4 %>% group_by(Population, Treatment) %>% mutate(relfit = estimates/mean(estimates)) %>% summarize(eye = var(relfit))

oppsel %>% group_by(Treatment) %>% summarize(med = median(eye), mini = min(eye), maxi=max(eye))

ggplot(aes(x=Population, y=eye, color=Treatment), data=oppsel)+geom_point()+theme_classic(base_size=16)

oppsel_checks <- seeds_work4 %>% group_by(Population,Treatment) %>% summarize(mean=mean(estimates)) 
```


Run an lmer for each population

```{r, warning=FALSE, message=FALSE}
#creating a function that extracts the emmeans for models
f <- function(df) as.data.frame(emmeans(lmer(estimates ~ Treatment +(1|Plant_ID), data = df), "Treatment"))
#applying this function to the dataset subset by population
v <- lapply(split(seeds_work4, seeds_work4$Population), f)

#creating a function that extracts the pairwise comparisons of emmeans for models
f2 <- function(df) as.data.frame(pairs(emmeans(lmer(estimates ~ Treatment +(1|Plant_ID), data = df), "Treatment")))
#applying this function to the dataset subset by population
v3 <- lapply(split(seeds_work4, seeds_work4$Population), f2)

#making the contrasts object more usable, adding significance column
v4 <-do.call(rbind.data.frame, v3)
v4$Pop<-rownames(v4)
rownames(v4)<-c()
v4$sig<-ifelse(v4$p.value<0.05, "*", "")

#making the emmeans object more usable
v2<-do.call(rbind.data.frame, v)
Pop=sort(as.factor(c(unique(seeds_work2$Population),unique(seeds_work2$Population))))
v2<-data.frame(v2, Pop, row.names = c())

#printing these tables
kable(v2)
kable(v4)

#plotting the pollen limitation results
ggplot(aes(x=Treatment, y=emmean), data=v2) + facet_wrap(~Pop) +geom_point()+geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=0.2) +ylab("Seed set") + theme_classic()+geom_text(aes(x=1.5, y=20, label=sig),data=v4, size=6 )

```

```{r}
#trying to work with size

#first on the population level

PL <- v2 %>%  pivot_wider(names_from=Treatment, values_from = emmean, values_fill = 0) %>% group_by(Pop) %>% mutate(C=sum(c), HP=sum(hp), PL=(HP-C)/HP)

size_field <- field %>% filter(FlowerSize !="NA") %>%  group_by(Population) %>% summarize(size = mean(FlowerSize))

PL_pop <- inner_join(size_field, PL, by=c("Population"="Pop")) %>% distinct(Population, .keep_all=TRUE)

ggplot(aes(x=size, y=log(PL+0.0001), label=Population), data=PL_pop)+geom_smooth(method="lm", se=FALSE)+geom_label_repel(min.segment.length = Inf)+geom_point()

model <- lm(PL~size, data=PL_pop)

cor(PL_pop$size, PL_pop$PL)

#now on the plant level

seeds_work4<- seeds_work2 %>% mutate(estimates=if_else(estimates<0, 0, estimates)) %>% mutate(estimates=if_else(estimates==333, 33, estimates)) %>% mutate(estimates=if_else(is.na(estimates), Seeds_counted, estimates)) %>% drop_na(estimates)

plant_PL2 <- seeds_work4 %>% group_by(ID, Treatment) %>% summarize(M = mean(estimates)) %>% pivot_wider(names_from = Treatment, values_from = M)%>% mutate(PL=(hp-c)/hp) %>% left_join(., field, by="ID") %>% filter(PL > -15) %>% filter(FlowerSize >1)


library(ggpubr)

##this plot##
#with colors

ggplot(aes(x=FlowerSize, y=PL), data=plant_PL2)+geom_point()+geom_smooth(method="lm", se=FALSE)+facet_wrap(~Population)+scale_y_continuous(limits=c(-15, 2))+theme_classic(base_size = 16)+theme(legend.position = "none")



plant_PL2$Population <- as.factor(plant_PL2$Population)
levels(plant_PL2$Population) <- c( "G3",  "G6" , "G7" , "G8",  "G9" , "I10" ,"I15", "I16", "I17", "I4" , "I7","G1"  )
model <- lm(FlowerSize~PL+Population,data=plant_PL2)

model<-lm(PL~FlowerSize*Population, data=plant_PL2)
hist(resid(model))
plot(predict(model),resid(model)) ; abline(h=0)


par(mfrow=c(1,2))
hist(resid(model))
plot(predict(model),resid(model)) ; abline(h=0)

linesPL <- coef(model)$Population
linesPL$Population <- rownames(linesPL)

ggplot(aes(x=FlowerSize, y=PL, color=Population, group=Population), data=plant_PL2)+geom_point()+facet_wrap(~Population)

plant_PL2 %>%
  group_by(Population) %>%
  summarize(COR=cor(PL,FlowerSize))

```

Figure 2
```{r}
#panel A:


a <- ggplot(aes(x=Treatment, y=emmean), data=v2) + facet_wrap(~Pop) +geom_point()+geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), width=0.2) +ylab("Seed set") + theme_classic(base_size = 16)+geom_text(aes(x=1.5, y=20, label=sig),data=v4, size=6 ) 

PL_pop$sig <- v4$sig
#panel B:
b <- ggplot(aes(x=size, y=PL, label=Population), data=PL_pop)+geom_smooth(method="lm", se=FALSE)+geom_point(aes(shape=sig), size=6)+geom_label_repel(min.segment.length = Inf)+theme_classic(base_size = 16)+scale_shape_manual(values=c(1,8))+theme(legend.position = "none")+xlab(expression(paste("Corolla area, (", mm^2 ,")")))+ylab("Pollen limitation")

+geom_errorbarh(aes(xmin=size-SE, xmax=size+SE), data=PL_pop)

ggarrange(a, b)

```



Working with selection.
First, estimating selection in the HP and C groups separately.

```{r}
f <- function(df) summary(lm((hp/mean(hp)) ~ scale(FlowerSize), data = df))
#applying this function to the dataset subset by population
v <- lapply(split(plant_PL2, plant_PL2$Population), f)

Pop=sort(as.factor(c(unique(plant_PL2$Population),unique(plant_PL2$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

##extracting the coefficients from each model
for (i in 1:12){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

#putting the results together with population labels
results<-cbind(results, Pop)
results<-results %>% tibble::rownames_to_column("Coefficient") 

#the only population with selection in the hp fruits is It15.
```

C fruits.

```{r}
f <- function(df) summary(lm((c/mean(c)) ~ scale(FlowerSize), data = df))
#applying this function to the dataset subset by population
v <- lapply(split(plant_PL2, plant_PL2$Population), f)

Pop=sort(as.factor(c(unique(plant_PL2$Population),unique(plant_PL2$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

##extracting the coefficients from each model
for (i in 1:12){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

#putting the results together with population labels
results<-cbind(results, Pop)
results<-results %>% tibble::rownames_to_column("Coefficient") 
```

Comparing the treatments
```{r}

plants_PL_long <- plant_PL2 %>% pivot_longer(cols=c("hp", "c"), names_to="Treatment")

all_pops_model <- lmer((value/mean(value))~scale(FlowerSize)*Treatment*Population+(1|Plant_ID), data = plants_PL_long)

f <- function(df) anova(lmer((value/mean(value)) ~ scale(FlowerSize)*Treatment+(1|Plant_ID), data = df))
#applying this function to the dataset subset by population
v <- lapply(split(plants_PL_long, plants_PL_long$Population), f)

Pop=sort(as.factor(c(unique(plants_PL_long$Population),unique(plants_PL_long$Population))))

results<-data.frame("Estimate"=numeric(), "Std. Error"=numeric(), "df"=numeric(), "t value"=numeric(), "Pr(>|t|)"=numeric())

##extracting the coefficients from each model
for (i in 1:12){
  z<-as.data.frame(v[[i]]$coefficients)
  results<-rbind(results,z)
}

#putting the results together with population labels
results<-cbind(results, Pop)
results<-results %>% tibble::rownames_to_column("Coefficient") 
```

